---
title: "Group 13 GLM - Ikea Dataset"
author: "Group 13"
date: "14/03/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadpackages, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(dplyr)
library(zoo)
```

The dataset contains information on IKEA furniture from Saudi Arabia, a column has been added indicating if the piece of furniture costs more than 1000 Riyals.



```{r dataset, echo=FALSE, eval=TRUE, warning=FALSE,message=FALSE}
data<- read.csv("dataset13.csv")

data<-mutate(data, thousandRiyals=ifelse(price>= 1000,"Yes", "No"))
head(data)
```

The summary of the data indicates that there is a lot of NA values in the height, depth and width columns.
 
``` {r summary, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
summary(data)
```

To solve this, I computed the mean of the depth, height and width by the category and replaced the NA's by this value shown below.

``` {r replace NA, echo=FALSE, eval=TRUE, warning=FALSE, message= FALSE}
cols<-c("depth","height","width")
data<-data%>%
  group_by(category)%>%
  mutate_at(cols, na.aggregate)
data
```







